---
title: "PSTAT 126 - Assignment 4"
subtitle: "Fall 2022"
author: "Due: Tuesday, October 25 at 11:59 pm on Canvas"
output:
  pdf_document:
    number_sections: true
---

*Note: **Submit both your `Rmd` and generated pdf file to Canvas.** Use the same indentation level as **Solution** markers to write your solutions. Improper indentation will break your document.*

```{r setup, include=FALSE}
# formats rendered outputs from inline R code
# Register an inline hook:
knitr::knit_hooks$set(inline = function(x) {
  x <- sprintf("%1.2f", x)
  paste(x, collapse = ", ")
})
```

```{r message=FALSE}
library(alr4)
library(ggplot2)
data(UN11)
```

1.  This problem uses the data set `UN11` from the `alr4` package.

    (a) Examine the figure generated by using `scatterplotMatrix` function for attributes (`fertility`, log(`ppgdp`), `pctUrban`), and comment on the marginal relationships.

        **Solution**:

```{r}
scatterplotMatrix(~fertility+log(ppgdp)+pctUrban, data= UN11)
```

Fertility and log(ppgdp) have a negative correlation, Fertility and pctUrban have a negative correlation, log(ppgdp) and pctUrban have a positive correlation.

    (b) Fit the two simple regressions `fertility` $\sim$ log(`ppgdp`) and `fertility` $\sim$ `pctUrban`, and verify that the slopes are significantly different from zero at any conventional level of significance.

        **Solution**:

```{r}
fit1 <- lm(fertility ~ log(ppgdp), data = UN11)
fit2 <- lm(fertility ~ pctUrban,data = UN11)
summary(fit1)
summary(fit2)
```

Observing the above summaries, Yes, the slopes are significantly different than zero at any conventional level.

    (c) Obtain the added-variable plots for both predictors. Based on the added-variable plots, does log(`ppgdp`) seem to be useful after adjusting for `pctUrban`, and similarly, does `pctUrban` seem to be useful after adjusting for log(`ppgdp`)?

        **Solution**:

```{r}
fit <- lm(fertility ~ log(ppgdp)+pctUrban,data = UN11)
avPlots(fit)
```

The av plot for log(ppgdp) after adjusting for pctUrban is useful as it maintains a steep slope, while pctUrban after adjusting for log(ppgdp) has a neutral slope and is not useful.

2.  Consider a multiple linear regression model with two continuous predictors:\
    $$
    Y_i = \beta_0 + \beta_1 x_{i1} + \beta_2 x_{i2} + \varepsilon_i, \qquad \varepsilon_i \stackrel{iid}{\sim} \mathcal{N}(0, \sigma^2).
    $$

    (a) Suppose that $x_{i1}$ and $x_{i2}$ are exactly related in that $x_{i1} = 2.2x_{i2}$ for all $i$. For example, $x_{i2}$ could be weight in kilograms and $x_{i1}$ weight in pounds for the $i$-th individual. Describe the appearance of the added variable plot for $x_{i2}$ after adjusting for $x_{i1}$.

        **Solution**:

Since xi2 is a linear function of xi1 and are exactly related, the residuals are zero and therefor the avPlot will look like a vertical line.

    b)  Suppose that $x_{i1}$ and $x_{i2}$ are not perfectly correlated, but that $Y_i = 3x_{i1}$, i.e. $Y_i$ is perfectly correlated with $x_{i1}$. Describe the added-variable plot for $x_{i2}$.

        **Solution**:
        

Since xi1 and xi2 are not perfectly correlated, the avPlot for xi2 would look like a straight line (no-slope or little-to-no-slope)

    c)  (**Bonus**): Simulate some data for each of the situations in parts a) and b) and create an added-variable plot to confirm you answers.

        **Solution**:
