---
title: "PSTAT 126 - Assignment 6"
subtitle: "Fall 2022"
author: "Due: Tuesday, November 15 at 11:59 pm"
output:
  pdf_document:
    number_sections: true
editor_options: 
  markdown: 
    wrap: 72
---

*Note: **Submit both your `Rmd` and generated pdf file to Gradescope.**
Use the same indentation level as **Solution** markers to write your
solutions. Improper indentation will break your document.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, message=FALSE}
library(tidyverse)
```

1.  Using the prostate data from the faraway package with lpsa (log
    prostate specific antigen) as response and lcavol (log cancer
    volume) as predictor, the fitted model is $$
    \text { lpsa }=1.507+0.719 \text { lcavol }
    $$

Provide an interpretation of the estimated coefficient for lcavol based
on the fact that both variables are log-transformed.

**Solution**:

Given lpsa and lcavol are both log-transformed, and with the lcavol
coefficient of 0.719, then the expected value of lpsa is changed by
100[(1+p)\^0.719-1]% where p is the % change in lcavol. For example, if
lcavol is changed by 10%, then lpsa changes by 100[(1+.10)\^0.719-1] =
7.09306%

2.  In a study of faculty salaries in a small college in the Midwest, a
    linear regression model was fit, giving the fitted mean function $$
    E(\text { Salary } \mid \text { Sex })=24697-3340 \text { Sex, }
    $$ where Sex equals 1 if the faculty member was female and 0 if
    male. The response Salary is measured in dollars (the data are from
    the 1970s).

    (a) Give a sentence that describes the meaning of the two estimated
        coefficients.

    **Solution**:

24697 is the expected salary for faculty members if male (0) and -3340
is the decrease in salary if the faculty member is a female (1).

(b) An alternative mean function fit to these data with an additional
    term, Years, the number of years employed at this college, gives the
    estimated mean function $$
      E(\text { Salary } \mid \text { Sex, Years })=18065+201 \text { Sex }+759 \text { Years } .
      $$ The important difference between these two mean functions is
    that the coefficient for Sex has changed signs. Provide an
    explanation as to how this could happen.

**Solution**:

Now, the salary for employees are not only considered by their sexuality
but also their employment time in the company. So, an explanation as to
why the coefficient has changed signs could be that the male faculty
members on average have been employed longer than female faculty
members. So if we were to control for the number of years employed, the
salary of female faculty members is actually higher than that of males.
Thus in the first model, we do not account for years employed and so it
seems that the salary of male faculty members is larger. Hence the
coefficient for sex has changed signs.

3.  This problem uses the data set cakes from the alr4 package, which
    contains the results of a baking experiment on n = 14 packaged cake
    mixes. The variables X1 and X2 data are the predictors representing
    baking time in minutes and baking temperature in degrees Fahrenheit,
    respectively. The response Y is a palatability score indicating
    quality of the cake.

```{r}
library(alr4)
```

(a) Fit the model\
    $$
      E(\mathrm{Y} \mid \mathrm{X} 1, \mathrm{X} 2)=\beta_0+\beta_1 \mathrm{X} 1+\beta_2 \mathrm{X} 2+\beta_{11} \mathrm{X} 1^2+\beta_{22} \mathrm{X} 2^2+\beta_{12} \mathrm{X} 1 \mathrm{X} 2
      $$ and verify that the p-values for the quadratic terms and the
    interaction are all less than 0.005.

**Solution**:

```{r}
data(cakes)
attach(cakes)
fit <- lm(Y~ X1+X2+I(X1^2)+I(X2^2)+I(X1*X2), data = cakes)
summary(fit)
anova(fit)
```

We can see by the summary and anova tables that the p-values are indeed
all less than 0.005.

(b) The cake experiment was carried out in two blocks of seven
    observations each. It is possible that the response might differ by
    block, due to differences in air temperature or humidity, for
    example. Add a main effect for the Block variable to model in part
    a), fit the model, and summarize results.

**Solution**:

```{r}
fit1 <- lm(Y~ X1+X2+I(X1^2)+I(X2^2)+I(X1*X2)+ block, data = cakes)
summary(fit1)
anova(fit1)
```

We can see from the table that the block variable is not significant.

4.  The data BGSall in the alr4 package contains information on n = 136
    children in the Berkeley Guidance study, including heights at ages 9
    and 18 (HT9 and HT18), and gender (Sex = 0 for male, 1 for female).
    Consider the regression of HT18 on HT9 and the grouping factor Sex.

```{r}
data('BGSall')
attach(BGSall)
```

(a) Draw the scatterplot of HT18 versus HT9, using a different symbol
    for males and females. Comment on the information in the graph about
    an appropriate mean function for these data.

**Solution**:

```{r}
HT18 <- BGSall$HT18
HT9 <- BGSall$HT9
sex <- BGSall$Sex

plot(HT9,HT18,type = 'n', main = 'Height Age 9 vs. Height Age 18')
points(HT9[Sex==0], HT18[Sex==0],col = 'blue', pch =15)
points(HT9[Sex==1], HT18[Sex==1],col = 'magenta', pch = 16)
legend('topleft',bty='n',col=c('blue','magenta'),c('Male','Female'),pch =c(15,16))
```

The appropriate mean function for this data would be fitting a parallel model. Here, we see that there is no interaction between gender and age groups.

(b) Obtain the appropriate test for a parallel regression model.

**Solution**:

```{r}
fit2 <- lm(HT18 ~ HT9 + factor(Sex), data = BGSall)
fit3 <- lm(HT18 ~ HT9*factor(Sex), data = BGSall)
anova(fit2,fit3)
```

The p-value is 0.08749, greater than 0.05; thus we accept the null hypothesis that the coefficient for interaction is 0 at a 5% confidence interval. Hence the parallel regression model is adequate. 

(c) Assuming the parallel regression model is adequate, estimate a 95%
    confidence interval for the difference between males and females.
    For the parallel regression model, this is the difference in the
    intercepts of the two groups.

**Solution**:

```{r}
summary(fit2)
```
Observing the summary, we see that The difference in the intercepts of the two groups (male and female) is 11.69584. SO, the estimated difference between males and females is 11.69584 and the standard error is 0.59036 and degree of freedom of residual is equal to 133.Now constructing a t-value for the test,

```{r}
abs(qt(0.05/2, 133))
```

```{r}
(11.69584 - 1.98 * 0.59036)
(11.69584 + 1.98 * 0.59036)
```
Hence, our 95% confidence interval between males and females is :
(10.52693, 12.86475)

5.  The data set infmort from the faraway package contains information
    on the mortality of infants for 105 nations. The variable mortality
    gives the number of deaths per 1000 live births, while income is the
    per capita income in US dollars and region indicates the geographic
    area of the nation. Consider the model $$
    E(\log (\text { mortality }) \mid \text { income,region })=\beta_0+\beta_1 \log (\text { income })+\beta_2 \text { region }+\beta_{12} \text { region } * \log (\text { income })
    $$
```{r}
library("faraway")
data("infmort")
```
    (a) State the null and alternative hypotheses for the overall F-test
        for this model. Perform the test and summarize results.

    **Solution**:

Null Hypothesis:
$$ H_0: \beta_1 = \beta_2 =\beta_{12}= 0 $$
Alternative hypothesis:
$$ H_1: \text {At least one of } \beta_{ai}\neq 0 $$
Perform the test:
```{r}
fit4 <- lm(log(mortality)~region+log(income)+log(income)*region, data = infmort)
summary(fit4)
```

From Summary, observe that F= 24.29 on 7 and 93 DF and p-value is 2.2e-16.
Therefore, we reject the null hypothesis since p-value is less than the 0.05, and
we can conclude that the fitted model is significant.


(b) Explain the practical meaning of the hypothesis
    $\mathcal{H}_0: \beta_{12}=\beta_2=0$ in the context of the above
    model.

**Solution**:
Of the above model,  $\mathcal{H}_0: \beta_{12}=\beta_2=0$ implies that the region has no impact on the relationship between income and mortality, so when given income and region, log(mortality) is independent of the region and interaction between the region and log(income).

## c).

(c) Perform a test for the hypothesis in part b) and summarize your
    results.

**Solution**:

```{r}
m1 <- lm(log(mortality)~log(income), data = infmort)
m2 <- lm(log(mortality)~log(income)+ region +log(income)*region, data = infmort)
anova(m1, m2)
```

We see that the p-value is 1.31e-05, which is much smaller than 0.05. Therefore, we reject the null hypothesis that beta12 and beta2 are 0. Therefore, the region and interaction between region and log(income) are significant in determining the log(mortality).

